---
title: "Homework 5"
author: "Elizabeth Jamison"
date: "2/12/2020"
output: html_document
---

## Working with Matrices, Lists, and Data Frames

### Question 1

Assign to the variable n_dims a single random integer between 3 and 10.

```{r}
n_dims <- sample(3:10, 1)
print(n_dims)
```

Create a vector of consecutive integers from 1 to n_dims^2

```{r}
vec1 <- c(1:n_dims^2)
print(vec1)
```


Use the sample function to randomly reshuffle these values.

```{r}
sample(vec1, replace = FALSE)
```

create a square matrix with these elements.

```{r}
mat1 <- matrix (vec1, nrow = n_dims, ncol = n_dims)
```

print out the matrix.
```{r}
print(mat1)
```

find a function in r to transpose the matrix.
```{r}
mat1 <- t(mat1)
```

print it out again and note how it has changed.
```{r}
print(mat1) # Rows now fill horizontally as opposed to vertically
```

calculate the sum and the mean of the elements in the first row and the last row.

```{r}
sum(mat1[1,])
mean(mat1[1,])
sum(mat1[n_dims,])
mean(mat1[n_dims,])
```

read about the eigen() function and use it on your matrix

```{r}
eigen(mat1)
```

look carefully at the elements of $values and $vectors. What kind of numbers are these?

* They are imaginary numbers

dig in with the typeof() function to figure out their type.

```{r}
typeof(eigen(mat1)) # list
typeof(2.699571e+02+0.000000e+00i) # complex- imaginary numbers
```

if have set your code up properly, you should be able to re-run it and create a matrix of different size because n_dims will change.
 
 * re-ran the code and it worked
 
### Question 2

Create a list with the following named elements:
  my_matrix, which is a 4 x 4 matrix filled with random uniform values
  my_logical which is a 100-element vector of TRUE or FALSE values. Do this efficiently by setting up a vector of random values and then applying an inequality to it.
  my_letters, which is a 26-element vector of all the lower-case letters in random order.

```{r}
# my_matrix
my_matrix <- matrix(nrow = 4, ncol = 4, data = runif(16))
print(my_matrix)

# my_logical
vec_log <- sample(1:100, 100)
print(vec_log)
my_logical <- vec_log > 50
print(my_logical)

# my_letters
my_letters <- sample(letters, replace = FALSE)
print(my_letters)

# my_list
my_list <- list(my_matrix, my_logical, my_letters)
print(my_list)

```

create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.

```{r}
new_list <- list(my_matrix[2,2], my_logical[2], my_letters[2])
print(new_list)
```

use the typeof() function to confirm the underlying data types of each component in this list

```{r}
typeof(new_list[[1]])
typeof(new_list[[2]])
typeof(new_list[[3]])
```


combine the underlying elements from the new list into a single atomic vector with the c() function.

```{r}
vec2 <- c(new_list[[1]], new_list[[2]], new_list[[3]])
print(vec2)
```

what is the data type of this vector?

```{r}
typeof(vec2) # character
```

### Question 3

Create a data frame with two variables (= columns) and 26 cases (= rows).
call the first variable my_unis and fill it with a random uniform value from 0 to 10
call the second variable my_letters and fill it with 26 capital letters in random order

```{r}
# first variable
my_unis <- runif(26, min=0, max=10)
print(my_unis)

# second variable
my_letters <- sample(LETTERS, 26, replace = FALSE)
print(my_letters)

# data frame
df <- data.frame(my_unis, my_letters)
print(df)

```


for the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA.

```{r}
df[sample(1:26,4),1] <- NA
print(df)
```

for the first variable, write a single line of R code to identify which rows have the missing values.

```{r}
# create a new data frame that contains only the rows of df that have NA values
id_na_values <- df[rowSums(is.na(df)) > 0,]
print(id_na_values)

```

for the second variable, sort it in alphabetical order

```{r}
df <- df[order(df[,2]),]
print(df)
```

calculate the column mean for the first variable.

```{r}
mean(df[,1], na.rm = TRUE)
print(df)

```

