---
title: "Homework 10"
author: "Elizabeth Jamison"
date: "4/2/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

## For loops and randomization tests

1. Using a for loop, write a function to calculate the number of zeroes in a numeric vector. Before entering the loop, set up a counter variable counter <- 0. Inside the loop, add 1 to counter each time you have a zero in the matrix. Finally, use return(counter) for the output.

```{r}
counter <- 0

# ------------------------------------------------
# FUNCTION num_zeros
# Description: calculates the number of zeros in a numeric vector
# Inputs: numeric vector
# Outputs: "counter" which counts the number of zeros in the numeric vector
##################################################
num_zeros <- function(numeric_vec = c(0,1,2,3,4,5)) {
  for (i in 1:length(numeric_vec)) { # start of for loop
    if(numeric_vec[i] == 0) { # start of if statement
      counter <- counter + 1
    } # end of if statement
  } # end of for loop

return(counter)

} # end of num_zeros
#-------------------------------------------------

# testing num_zeros
num_zeros() # default

vec1 <- sample(0:10, 20, replace = TRUE)
print(vec1)
num_zeros(vec1)

```


2. Use subsetting instead of a loop to rewrite the function as a single line of code.

```{r}
# ------------------------------------------------
# FUNCTION num_zeros_sub
# Description: counts the number of zeros in a numeric vector using subsetting
# Inputs: numeric vector
# Outputs: "counter" which counts the number of zeros in the numeric vector
##################################################
num_zeros_sub <- function(numeric_vec = c(0,1,2,3,4,5)) {
  
  counter <- length(numeric_vec[numeric_vec == 0]) 
  
  return(counter)

} # end of num_zeros_sub
#-------------------------------------------------

# testing num_zeros_sub
num_zeros_sub() # default

vec1 <- sample(0:10, 20, replace = TRUE)
print(vec1)
num_zeros(vec1)

```


3. Write a function that takes as input two integers representing the number of rows and columns in a matrix. The output is a matrix of these dimensions in which each element is the product of the row number x the column number.

```{r}
mat1_nrow <- 4
mat1_ncol <- 5
mat1 <- matrix(nrow = mat1_nrow, 
               ncol = mat1_ncol, 
               data = sample(0:10, mat1_nrow*mat1_ncol, replace = TRUE))
print(mat1)

# ------------------------------------------------
# FUNCTION prod_matrix
# Description: creates a matrix in which each element is a product of the row number and column number
# Inputs: two integers representing the number of rows and columns in the matrix
# Outputs: matrix with dimensions corresponding to the integers
##################################################
prod_matrix <- function(num_row = 5, num_col = 4) {
  mat1 <- matrix(nrow = num_row, ncol = num_col)
  for (i in 1:num_row) { # start of outer loop
    for (j in 1:num_col) { # start of inner loop
      mat1[i,j] <- i*j
    } # end of inner loop
  } # end of outer loop

return(mat1)

} # end of prod_matrix
#-------------------------------------------------

# testing prod_matrix

prod_matrix()

prod_matrix(num_row = 6, num_col = 7)

```


4. Use the code from the upcoming April 2nd lecture (Randomization Tests) to design and conduct a randomization test for some of your own data. You will need to modify the functions that read in the data, calculate the metric, and randomize the data. Once those are set up, the program should run correctly calling your new functions. Also, to make your analysis fully repeatable, make sure you set the random number seed at the beginning (use either set.seed() in base R, or char2seed in the TeachingDemos package

The data I choose to use are the number of trees/hectare for 49 pitch pine barren stands. 40 of these stands are infested with southern pine beetle and 9 of these stands are uninfested. ID = stand number, x = status (infested or uninfested), y = number of trees/hectare (TPHA).

```{r}
# Preliminaties -----------------------------------------------------------

library(ggplot2)
require(ggplot2)
library(TeachingDemos)
require(TeachingDemos)

set.seed(100)

z <- read.csv("C:/Users/eliza/Desktop/JamisonBio381/TPHAdata_RandomizationTest_HW10.csv")
print(z)

# Building functions ------------------------------------------------------

# ------------------------------------------------
# FUNCTION read_data
# Description: read in (or generate) data set for analysis
# Inputs: file name (or nothing, as in this demo)
# Outputs: 3 column data frame of observed data (ID, x, y)
##################################################
read_data <- function(z = NULL) {
  if (is.null(z)) {
    x <- 1:20
    y_obs <- x_obs + 10*rnorm(20)
    df <- data.frame(ID=seq_along(x_obs),
                     x_obs,
                     y_obs)
  }
  
  df <- read.table(file = z,
                   header = TRUE,
                   sep = ",",
                   stringsAsFactors = FALSE)
  
return(df)

} # end of read_data
#-------------------------------------------------
#read_data()

# ------------------------------------------------
# FUNCTION get_metric
# Description: calculate metric for randomization test
# Inputs: 2-column data frame
# Outputs: difference in means between two groups
##################################################
get_metric <- function(z = NULL) {
  if(is.null(z)) {
    x <- c(rep("INF", 10), rep("UN", 10))
    y <- 10*rnorm(20)
    z <- data.frame(ID = seq_along(x),
                    x,
                    y)
  }
 
  diff_mean <- abs(mean(z[z$x == "INF",3]) - mean(z[z$x == "UN", 3]))

  return(diff_mean)

} # end of get_metric
#-------------------------------------------------
#get_metric()

# ------------------------------------------------
# FUNCTION shuffle_data
# Description: randomize data for a regression analysis
# Inputs: 3 column data frame (ID, xvar, yvar)
# Outputs: 3 column data frame(ID, xvar, yvar)
##################################################
shuffle_data <- function(z = NULL){
  if(is.null(z)) {
    x <- c(rep("a", 10), rep("b", 10))
    y <- 10*rnorm(20)
    z <- data.frame(ID = seq_along(x_obs),
                    x,
                    y)
  }
  
  sam <- sample(z$y)
  z$y <- sam

return(z)

} # end of shuffle_data
#-------------------------------------------------
#shuffle_data()

# ------------------------------------------------
# FUNCTION get_pval
# Description: calculate p value from simulation
# Inputs: list of observed metric and vector of simulated metrics
# Outputs: lower and upper tail probability values
##################################################
get_pval <- function(z = NULL) {
  if(is.null(z)) {
    z <- list(rnorm(1), rnorm(1000)) # z[[2]] contains 1000 simulated slopes
  }                                 # z[[1]] contains observed slope
  
  p_lower <- mean(z[[2]] <= z[[1]])
  p_upper <- mean(z[[2]] >= z[[1]])
  

return(c(p_lower, p_upper))

} # end of get_pval
#-------------------------------------------------
#get_pval()

# ------------------------------------------------
# FUNCTION plot_ran_test
# Description: create a ggplot of histogram of simulated values
# Inputs: list of observed metrics and vector simulated metrics
# Outputs: saved ggplot graph
##################################################
plot_ran_test <- function(z = NULL) {
  if (is.null(z)) {
    z <- list(rnorm(1), rnorm(1000))
  }
  
  df <- data.frame(ID = seq_along(z[[2]]), sim_x = z[[2]])
  p1 <- ggplot(data = df, mapping = aes(x = sim_x))
  p1 + geom_histogram(mapping = aes(fill = I("yellow"), color = I("black"))) +
    geom_vline(aes(xintercept = z[[1]], col="blue"))


return(p1 + geom_histogram(mapping = aes(fill = I("yellow"), color = I("black"))) +
         geom_vline(aes(xintercept = z[[1]], col="blue"))) # it would not print the graph unless I included this code in "return"

} # end of plot_ran_test
#-------------------------------------------------
#plot_ran_test()

# END OF FUNCTIONS

# --------------------------------------------------

# Assembling functions ---------------------------------

n_sim <- 1000 # number of simulated data sets
x_sim <- rep(NA, n_sim) # set up empty vector for simulated difference between means

z <- read_data("TPHAdata_RandomizationTest_HW10.csv") # read in data
print(z) # checking data


x_obs <- get_metric(z) # get difference between means of observed data
#print(x_obs)

for (i in seq_len(n_sim)) {
  x_sim[i] <- get_metric(shuffle_data(z)) # get difference between means of simulated data
}
#print(x_sim)

diff_means <- list(x_obs, x_sim)

get_pval(diff_means)

plot_ran_test(diff_means)

```


5. For comparison, calculate in R the standard statistical analysis you would use with these data. How does the p-value compare for the standard test versus the p value you estimated from your randomization test? If the p values seem very different, run the program again with a different starting seed (and/or increase the number of replications in your randomization test). If there are persistent differences in the p value of the standard test versus your randomization, what do you think is responsible for this difference?

```{r}
# Wilcoxon rank-sum test- a non parametric test (data is not normally distributed)
    # Is a randomly selected value from one population likely to be greater or less
    # than a randomly selected vale from another population?
wilcox.test(z$y ~ z$x, exact = FALSE, na.rm = TRUE)
get_pval(diff_means)

```

The p value from the Wilcoxon rank-sum test is close to, but not exactly the same as, the lower p value. This may be because the data is not normally distributed (see Shapiro-Wilk normality test below) or because there were more infested stands (40) than uninfested stands (9). 

```{r}
# Shapiro-Wilk normality test
# p > 0.05: there is no significant difference between data and a normal distribution

shapiro.test(z$y)

```

