---
title: "Homework 4"
author: "Elizabeth Jamison"
date: "2/5/2020"
output: html_document
---

## Question 1

Assign variables to their values:

```{r}
x <- 1.1
a <- 2.2
b <- 3.3
```

Assign an expression and store as z:

```{r}
z <- x^a^b 
z <- (x^a)^b
z <- 3*x^3 + 2*x^2 + 1
print(z)
```


## Question 2

Use the rep() and seq() functions to create these vectors:

```{r}
vec_a <- c(seq(from = 1, to = 8, by = 1), seq(from = 7, to = 1, by = -1))
print(vec_a)

vec_b <- c(1, rep.int(2, 2), rep.int(3,3), rep.int(4, 4), rep.int(5, 5))
print(vec_b)

vec_c <- c(5, rep.int(4, 2), rep.int(3,3), rep.int(2, 4), rep.int(1, 5))
print(vec_c)
```

## Question 3

Create a vector of 2 random uniform numbers:

```{r}
coord <- runif(2)
print(coord)
```

Convert these numbers into polar coordinates:

```{r}
r <- sqrt((coord[1])^2+(coord[2])^2)
print(r)

theta <- atan(coord[2]/coord[1])
print(theta)

pcord <- c(r,theta)
print(pcord)

```

## Question 4

Create initial vector:

```{r}
queue <- c("sheep", "fox", "owl", "ant")
print(queue)
```

Updating queue:

```{r}
queue_a <- c(queue, "serpent")
print(queue_a)

queue_b <- queue_a[-1]
print(queue_b)

queue_c <- c("donkey", queue_b)
print(queue_c)

queue_d <- queue_c[-5]
print(queue_d)

queue_e <- c(queue_d[-3])
print(queue_e)

queue_f <- c(queue_e[1:2], "aphid", queue_e[3])
print(queue_f)

which(queue_f == "aphid") # determin position of aphid

```

## Question 5

Create a vector of all of the integers from 1 to 100 that are not divisible by 2, 3, or 7

```{r}
vec <- Filter(function(i) 
  { all(i %% c(2,3,7) != 0) }, 
  seq(100))
print(vec)
```





