---
title: "Homework 12"
author: "Elizabeth Jamison"
date: "5/1/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

1. For this exerise, use your newly-developed ggplot chops to create some nice graphs from your own data (If you do not have a good data frame to use for graphics, use one of the many built-in data frames from R (other than mpg, which we are using in class)). Experiment with different themes, theme base sizes, aesthetics, mappings, and faceting. When you are finished, try exporting them to high quality pdfs, jpgs, eps files, or other formats that you would use for submission to a journal.


```{r}

# Preliminaries -----------------------------------------------------------

# LOAD LIBRARIES
library(ggplot2)
require(ggplot2)
library(patchwork)
require(patchwork)
library(scales)
require(scales)

# LOAD DATA
data <- read.csv("StandData_LongIsland.csv")

# GROUP DATA

# EJ, 2019
EJ <- data[data$DATA == "EJ",]

# MH, 2016
MH <- data[data$DATA == "MH",]

###############################################################

# BOX PLOTS ---------------------------------------------------------------

# basal area per ha: all stands

ba_boxplot <- ggplot(data = data, aes(x = STATUS, y = BA_HA)) +
  geom_boxplot(outlier.colour = "red") +
  labs(title = "Basal area per ha of infested and uninfested stands", 
       x = "Infestation status", 
       y = "Basal area per ha (m^2/ha)",
       caption = "Data: EJ, 2019 and MH, 2016") +
  scale_x_discrete(labels=c("INF" = "Infested", "UN" = "Uninfested")) +
  theme_bw(base_size = 13) +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.y = element_text(vjust = 3))
  
print(ba_boxplot)

# save box plot as png

ggsave("ba_boxplot.png")


# trees per ha: all stands
tpha_boxplot <- ggplot(data = data, aes(x = STATUS, y = TPHA)) +
  geom_boxplot(outlier.colour = "red") +
  labs(title = "Trees per ha of infested and uninfested stands",
       x = "Infestation status", 
       y = "Trees per ha",
       caption = "Data: EJ, 2019 and MH, 2016") +
  scale_x_discrete(labels=c("INF" = "Infested", "UN" = "Uninfested")) +
  theme_bw(base_size = 13) +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.y = element_text(vjust = 3))

print(tpha_boxplot)

# save box plot as pdf

ggsave("tpha_boxplot.pdf")



# proportion pitch HWMX: all stands
piri_boxplot <- ggplot(data = data, aes(x = STATUS, y = PROP_PIRI)) +
  geom_boxplot(outlier.colour = "red") +
  labs(title = "Proportion pitch pine of infested and uninfested stands",
       x = "Infestation status", 
       y = "Proportion pitch pine",
       caption = "Data: EJ, 2019 and MH, 2016") +
  scale_x_discrete(labels=c("INF" = "Infested", "UN" = "Uninfested")) +
  theme_bw(base_size = 13) +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.y = element_text(vjust = 3))

print(piri_boxplot)

# QMD: all data
QMD_boxplot <- ggplot(data = data, aes(x = STATUS, y = QMD)) +
  geom_boxplot(outlier.colour = "red") +
  labs(title = "QMD of infested and uninfested stands",
       x = "Infestation status", 
       y = "Quadratic mean diameter (cm)",
       caption = "Data: EJ, 2019 and MH, 2016") +
  scale_x_discrete(labels=c("INF" = "Infested", "UN" = "Uninfested")) +
  theme_bw(base_size = 13) +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.y = element_text(vjust = 3))

print(QMD_boxplot)

# SDI: EJ, 2019
SDI_boxplot <- ggplot(data = EJ, aes(x = STATUS, y = SDI)) +
  geom_boxplot(outlier.colour = "red") +
  labs(title = "SDI of infested and uninfested stands",
       x = "Infestation status", 
       y = "Stand density index",
       caption = "Data: EJ, 2019") +
  scale_x_discrete(labels=c("INF" = "Infested", "UN" = "Uninfested")) +
  theme_bw(base_size = 13) +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.y = element_text(vjust = 2))

print(SDI_boxplot) 

# relative density: EJ, 2019
RD_boxplot <- ggplot(data = EJ, aes(x = STATUS, y = REL_DEN)) +
  geom_boxplot(outlier.colour = "red") +
  labs(title = "Relative density of infested and uninfested stands",
       x = "Infestation status", 
       y = "Relative density",
       caption = "Data: EJ, 2019") +
  scale_x_discrete(labels=c("INF" = "Infested", "UN" = "Uninfested")) +
  theme_bw(base_size = 13) +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.y = element_text(vjust = 2))

print(RD_boxplot) 

###############################################################

```


```{r}

# PATCHWORK ---------------------------------------------------------------

###############################################################

# SDI and RELATIVE DENSITY

# plots for patchwork -----------------------------------------------------

# SDI: EJ, 2019
SDI_boxplot_PW <- ggplot(data = EJ, aes(x = STATUS, y = SDI)) +
  geom_boxplot(outlier.colour = "red") +
  labs(#title = "SDI of infested and uninfested stands",
       x = "Infestation status", 
       y = "Stand density index",
       caption = "Data: EJ, 2019") +
  scale_x_discrete(labels=c("INF" = "Infested", "UN" = "Uninfested")) +
  theme_bw(base_size = 13) +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.y = element_text(vjust = 2))

print(SDI_boxplot_PW) 

# relative density: EJ, 2019
RD_boxplot_PW <- ggplot(data = EJ, aes(x = STATUS, y = REL_DEN)) +
  geom_boxplot(outlier.colour = "red") +
  labs(#title = "Relative density of infested and uninfested stands",
       x = "Infestation status", 
       y = "Relative density",
       caption = "Data: EJ, 2019") +
  scale_x_discrete(labels=c("INF" = "Infested", "UN" = "Uninfested")) +
  theme_bw(base_size = 13) +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.y = element_text(vjust = 2))

print(RD_boxplot_PW)

# SDI + relative density --------------------------------------------------

SDI_RD_boxplots <- SDI_boxplot_PW + RD_boxplot_PW + 
  plot_annotation(title = "SDI and relative density of infested and uninfested stands", 
                  theme = theme(plot.title = element_text(hjust = 0.5)))

print(SDI_RD_boxplots)

###############################################################

# BASAL AREA, TPHA, PROPORTION HWMX, AND QMD

# plots for patchwork -----------------------------------------------------

# basal area per ha: all stands

ba_boxplot_PW <- ggplot(data = data, aes(x = STATUS, y = BA_HA)) +
  geom_boxplot(outlier.colour = "red") +
  labs(title = "Basal area per ha", 
       x = "Infestation status", 
       y = "Basal area per ha (m^2/ha)",
       caption = "Data: EJ, 2019 and MH, 2016") +
  scale_x_discrete(labels=c("INF" = "Infested", "UN" = "Uninfested")) +
  theme_bw(base_size = 9) +
  theme(axis.text.x = element_text(size = 9),
        axis.text.y = element_text(size = 9),
        axis.title.y = element_text(vjust = 3))

print(ba_boxplot_PW) 

# trees per ha: all stands
tpha_boxplot_PW <- ggplot(data = data, aes(x = STATUS, y = TPHA)) +
  geom_boxplot(outlier.colour = "red") +
  labs(title = "Trees per ha",
       x = "Infestation status", 
       y = "Trees per ha",
       caption = "Data: EJ, 2019 and MH, 2016") +
  scale_x_discrete(labels=c("INF" = "Infested", "UN" = "Uninfested")) +
  theme_bw(base_size = 9) +
  theme(axis.text.x = element_text(size = 9),
        axis.text.y = element_text(size = 9),
       axis.title.y = element_text(vjust = 3))

print(tpha_boxplot_PW) 


# proportion pitch HWMX: all stands
piri_boxplot_PW <- ggplot(data = data, aes(x = STATUS, y = BA_HA)) +
  geom_boxplot(outlier.colour = "red") +
  labs(title = "Proportion pitch HWMX",
       x = "Infestation status", 
       y = "Proportion pitch HWMX",
       caption = "Data: EJ, 2019 and MH, 2016") +
  scale_x_discrete(labels=c("INF" = "Infested", "UN" = "Uninfested")) +
  theme_bw(base_size = 9) +
  theme(axis.text.x = element_text(size = 9),
        axis.text.y = element_text(size = 9),
        axis.title.y = element_text(vjust = 3))

print(piri_boxplot_PW)

# QMD: all data
QMD_boxplot_PW <- ggplot(data = data, aes(x = STATUS, y = QMD)) +
  geom_boxplot(outlier.colour = "red") +
  labs(title = "QMD",
       x = "Infestation status", 
       y = "Quadratic mean diameter (cm)",
       caption = "Data: EJ, 2019 and MH, 2016") +
  scale_x_discrete(labels=c("INF" = "Infested", "UN" = "Uninfested")) +
  theme_bw(base_size = 9) +
  theme(axis.text.x = element_text(size = 9),
        axis.text.y = element_text(size = 9),
        axis.title.y = element_text(vjust = 3))

print(QMD_boxplot_PW)

# basal area + tpha + proportion HWMX + QMD
(ba_boxplot_PW + tpha_boxplot_PW) / (QMD_boxplot_PW + piri_boxplot_PW)+
  plot_annotation(title = "Stand metrics of infested and uninfested stands", 
                  theme = theme(plot.title = element_text(size = 13, hjust = 0.5)))

###############################################################

```


```{r}
# MULTIPLE BOX PLOTS ---------------------------------

# Box Plots: STATUS and COVER ----------------------------------------------

box_plot_colors <- c("lightcoral", "palegreen2")
box_plot_colors2 <- c("lightcoral", "palegreen2", "lightcoral", "palegreen2")
show_col(box_plot_colors)

# Basal area

ba_status_cover <- ggplot(data = data, aes(x = COVER, y = BA_HA, by = STATUS, fill = STATUS)) +
  geom_boxplot(outlier.color =  "black", 
               outlier.size = 2, 
               outlier.shape = "square") +
  geom_point(color = "grey30", 
            position = position_jitter(width = 0.1, height = 0),
            size = 2) +
  labs(title = "Basal area by stand classification and infestation status",
       x = "Stand classification",
       y = "Basal area per ha (m2/ha)",
       fill = "Infestation status") +
  scale_x_discrete(labels=c("HWMX" = "Oak-pine", "PINE" = "Pine")) +
  scale_fill_manual(labels = c("Infested", "Uninfested"), values = box_plot_colors) +
  theme_bw(base_size = 13) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
  #theme(axis.text.x = element_text(size = 12),
  #      axis.text.y = element_text(size = 12),
  #      axis.title.y = element_text(size = 12),
  #      axis.title.y = element_text(size = 12))


print(ba_status_cover)

# Trees per ha

tpha_status_cover <- ggplot(data = data, aes(x = COVER, y = TPHA, by = STATUS, fill = STATUS)) +
  geom_boxplot(outlier.color =  "black", 
               outlier.size = 2, 
               outlier.shape = "square") +
  geom_point(color = "grey30", 
             position = position_jitter(width = 0.1, height = 0),
             size = 2) +
  labs(title = "Trees per ha by stand classification and infestation status",
       x = "Stand classification",
       y = "Trees per ha",
       fill = "Infestation status") +
  scale_x_discrete(labels=c("HWMX" = "Oak-pine", "PINE" = "Pine")) +
  scale_fill_manual(labels = c("Infested", "Uninfested"), values = box_plot_colors) +
  theme_bw(base_size = 13) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
#theme(axis.text.x = element_text(size = 12),
#      axis.text.y = element_text(size = 12),
#      axis.title.y = element_text(size = 12),
#      axis.title.y = element_text(size = 12))


print(tpha_status_cover)

# QMD

QMD_status_cover <- ggplot(data = data, aes(x = COVER, y = QMD, by = STATUS, fill = STATUS)) +
  geom_boxplot(outlier.color =  "black", 
               outlier.size = 2, 
               outlier.shape = "square") +
  geom_point(color = "grey30", 
             position = position_jitter(width = 0.1, height = 0),
             size = 2) +
  labs(title = "Trees per ha by stand classification and infestation status",
       x = "Stand classification",
       y = "QMD",
       fill = "Infestation status") +
  scale_x_discrete(labels=c("HWMX" = "Oak-pine", "PINE" = "Pine")) +
  scale_fill_manual(labels = c("Infested", "Uninfested"), values = box_plot_colors) +
  theme_bw(base_size = 13) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
#theme(axis.text.x = element_text(size = 12),
#      axis.text.y = element_text(size = 12),
#      axis.title.y = element_text(size = 12),
#      axis.title.y = element_text(size = 12))


print(QMD_status_cover)

# relative density

RD_status_cover <- ggplot(data = EJ, aes(x = COVER, y = REL_DEN, by = STATUS, fill = STATUS)) +
  geom_boxplot(outlier.color =  "black", 
               outlier.size = 2, 
               outlier.shape = "square") +
  geom_point(color = "grey30", 
             position = position_jitter(width = 0.1, height = 0),
             size = 2) +
  labs(title = "Relative density by stand classification and infestation status",
       x = "Stand classification",
       y = "Relative density",
       fill = "Infestation status") +
  scale_x_discrete(labels=c("HWMX" = "Oak-pine", "PINE" = "Pine")) +
  scale_fill_manual(labels = c("Infested", "Uninfested"), values = box_plot_colors) +
  theme_bw(base_size = 13) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

#theme(axis.text.x = element_text(size = 12),
#      axis.text.y = element_text(size = 12),
#      axis.title.y = element_text(size = 12),
#      axis.title.y = element_text(size = 12))


print(RD_status_cover)
###############################################################

```

```{r}

# SCATTER PLOTS -----------------------------------------------------------

# Status ------------------------------------------------------------------

# basal area and TPHA

ba_tpha_scatterplot <- ggplot(data = data, mapping = aes(x = TPHA, y = BA_HA, color = STATUS)) +
  geom_point(size = 3) + 
  geom_smooth(method = "lm") +
  labs(title = "Trees per ha and basal area for uninfested and infested stands",
       subtitle = "Data: EJ, 2019 and MH, 2016",
       x = "Trees per ha",
       y = "Basal area per ha",
       color = "Infestation status") +
  theme_bw(base_size = 12) + 
  theme(axis.text.x = element_text(size = 11),
        axis.text.y = element_text(size = 11),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12, vjust = 3)) +
  scale_color_manual(labels = c("Infested", "Uninfested"), values = c("tomato2", "forestgreen"))

print(ba_tpha_scatterplot)

# QMD and TPHA

QMD_tpha_scatterplot <- ggplot(data = data, mapping = aes(x = TPHA, y = QMD, color = STATUS)) +
  geom_point(size = 3) + 
  geom_smooth(method = "lm") +
  labs(title = "Trees per ha and QMD for uninfested and infested stands",
       subtitle = "Data: EJ, 2019 and MH, 2016",
       x = "Trees per ha",
       y = "Quadratic mean diameter (cm)",
       color = "Infestation status") +
  theme_bw(base_size = 12) + 
  theme(axis.text.x = element_text(size = 11),
        axis.text.y = element_text(size = 11),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12, vjust = 3)) +
  scale_color_manual(labels = c("Infested", "Uninfested"), values = c("tomato2", "forestgreen"))

print(QMD_tpha_scatterplot)


# Cover -------------------------------------------------------------------

# Proportion HWMX and trees per ha
tpha_piri_scatterplot <- ggplot(data = data, mapping = aes(x = TPHA, y = PROP_PIRI, color = COVER)) +
  geom_point(size = 3) + 
  geom_smooth(method = "lm") +
  labs(title = "Trees per ha and proportion pitch pine by stand classification",
       subtitle = "Data: EJ, 2019 and MH, 2016",
       x = "Trees per ha",
       y = "Proportion pitch pine",
       color = "Stand classification") +
  theme_bw(base_size = 10) + 
  theme(axis.text.x = element_text(size = 11),
        axis.text.y = element_text(size = 11),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12, vjust = 3)) +
  scale_color_manual(labels = c("Oak-pine", "Pine"), values = c("mediumorchid", "seagreen"))

print(tpha_piri_scatterplot)

# QMD by tpha
tpha_QMD_scatterplot <- ggplot(data = data, mapping = aes(x = TPHA, y = QMD, color = COVER)) +
  geom_point(size = 3) + 
  geom_smooth(method = "lm") +
  labs(title = "Trees per ha and QMD by stand classification",
       subtitle = "Data: EJ, 2019 and MH, 2016",
       x = "Trees per ha",
       y = "Quadratic mean diameter (cm)",
       color = "Stand classification") +
  theme_bw(base_size = 12) + 
  theme(axis.text.x = element_text(size = 11),
        axis.text.y = element_text(size = 11),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12, vjust = 3)) +
  scale_color_manual(labels = c("Oak-Pine", "Pine"), values = c("mediumorchid", "seagreen"))

print(tpha_QMD_scatterplot)

# QMD by basal area
ba_QMD_scatterplot <- ggplot(data = data, mapping = aes(x = BA_HA, y = QMD, color = COVER)) +
  geom_point(size = 3) + 
  geom_smooth(method = "lm") +
  labs(title = "Basal area and QMD by stand classification",
       subtitle = "Data: EJ, 2019 and MH, 2016",
       x = "Basal area (m^2/ha)",
       y = "Quadratic mean diameter (cm)",
       color = "Stand classification") +
  theme_bw(base_size = 12) + 
  theme(axis.text.x = element_text(size = 11),
        axis.text.y = element_text(size = 11),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12, vjust = 3)) +
  scale_color_manual(labels = c("Oak-Pine", "Pine"), values = c("mediumorchid", "seagreen"))

print(ba_QMD_scatterplot)

###############################################################

# I was receiving error 1033 when trying to knit my document
# I found this code online that allowed me to knit the document after I ran it
my_locale <- Sys.getlocale("LC_ALL")
Sys.setlocale("LC_ALL", my_locale)

```

